---
title: "Homework week 4"
author: "Sigrid Agersnap Bom Nielsen"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# loading packages
pacman::p_load(tidyverse)
```


For this assignment, please answer the following three questions showing your code and results and then take a particular screenshot of your working environment. You can upload here (to Brightspace) the code and the URL to the screenshot OR submit everything to Github and submit here a single link to your repository.


# 1) Use R to figure out how many elements in the vector below are greater than 2.
```{r}
# vector
rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)

# solution
sum(rooms > 2, na.rm = T)

```
Answer: There are 8 elements of the vector which are greater than 2.
na.rm = T disregards the NA's of the vector

# 2) What type of data is in the 'rooms' vector?

Numberical entities and N/A's

# 3) What is the result of running the median() function on the above 'rooms' vector?

```{r}
# running the median function without doing something about the NA's
median(rooms)

# the output says NA because it cannot calculate the median of NA's

# if I write
median(rooms, na.rm = T)

# The output is 1.5, because it only considers the numerical proporties of the vector.
```


# 4) Submit the following image to Github:
Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing
a) the line of code you used to create the object,
b) the 'interviews' object in the Environment, and
c) the file structure of your R project in the bottom right "Files" pane.
Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.

```{r}
############### One way of doing it
# download the file directly from the link into R
# I call the dataframe df for esay access
df <- read_csv("https://ndownloader.figshare.com/files/11492171")

################ Another way of doing it
# download the file to my computer
download.file("https://ndownloader.figshare.com/files/11492171", "data/SAFI_clean.csv", mode = "wb")

# then load it into R
interviews <- read_csv("data/SAFI_clean.csv")
```

About the file structure:
I like to put my things in folders named by date. So next to the project file, I have a folder called '2021_09_24_HW4'. In that folder I have my Rmarkdown file and a data folder. Inside the data folder is the SAFI_clean.csv. 

Link to screen shot:
